Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-02 fib2.tex
polishing text.

These manipulations encode the data of the unitary modular tensor category, and can be used to represent the state of a system of anyons as a linear combination of fusion trees. For the Fibonacci anyons, the non-trivial $R$ and $F$ moves are given by
\begin{align}
	R_{\mathbb{I}}^{\tau\tau}&= \e^{\frac{-4\pi i}{5}}\qquad R_\tau^{\tau\tau}= \e^{\frac{3\pi i}{5}}\\
	F_{\tau}^{\tau\tau\tau} &= \begin{pmatrix}\varphi^{-1}&\varphi^{-\frac{1}{2}}\\\varphi^{-\frac{1}{2}}&-\varphi^{-1}\end{pmatrix} \,,
\end{align}
where in the last expression, $\varphi=\frac{1+\sqrt{5}}{2}$ is the golden ratio, and the matrix is given in a basis labelled $(\mathbb{I},\tau)$.

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate connected components 
of average size $O(\log(n))$ and variance $O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected components correspond to sets of anyons that could not have interacted at any 
point in their history. 
Although normally anyons can interact by braiding without ever coming 
into contact with one another \dude{um um um}, in this case charge conservation 
implies that each distinct component must have trivial total charge, and so all such interactions are trivial.
This allows us to simulate the braiding processes within each component 
separately. In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over components, and we can compute each factor separately. Since each 
component has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are rare 
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

In \Fref{f:syndrome} we show two adjacent sites
involved in a syndrome measurement.
%This will be the total charge on each site, ie. either $\tau$ or $\mathbb{I}.$
The error process has resulted in two pair
creation events along the edge joining these sites.
\dude{i'd like to write out the equations for each of
these 4 steps, but we'd need to keep it tight somehow.}
First, (a) we show the state as a tensor product of two pairs $\tau\tau\otimes\tau\tau$.
Then, (b) these two pairs are embedded into a single larger fusion
tree $\tau\tau\tau\tau$.
This is continuosly deformable into the picture in (c).
It is now clear the two braid operations required so
that the observables are represented on the fusion tree.
The general algorithm and datastructures used
for this proceedure will be discussed elsewhere.

@@

\stf{Cite~\cite{Pfeifer2010} for braiding tricks}

@@

There are several interesting avenues for further research. 
Although it is not completely obvious how to do so, applying these methods to more realistic models such as concrete microscopic spin models or models with non-topological features would give direct insight into practical error rates needed for topological quantum memories in non-Abelian systems. 
\dude{This doesn't make much sense to me:} Another potential direction is to incorporate tensor network methods for simulating anyons~\cite{Pfeifer2012, Singh2014} to improve the speed and scale of the simulations. 
Finally, perhaps the most important open question is whether 
\emph{fault-tolerant} decoders can be designed for nonabelian anyon codes. 
\dude{Another important question is wether this error
correction process can be mapped to a classical stat mech problem, like RBIM.}

>>>>

These manipulations encode the data of the unitary modular tensor category, and can be used to represent the state of a system of anyons as a linear combination of fusion trees. For the Fibonacci anyons, the non-trivial $R$ and $F$ moves are given by
\begin{align}
	R_{\mathbb{I}}^{\tau\tau}&= \e^{\frac{-4\pi i}{5}}\qquad R_\tau^{\tau\tau}= \e^{\frac{3\pi i}{5}}\\
	F_{\tau}^{\tau\tau\tau} &= \begin{pmatrix}\varphi^{-1}&\varphi^{-\frac{1}{2}}\\\varphi^{-\frac{1}{2}}&-\varphi^{-1}\end{pmatrix} \,,
\end{align}
where the matrix is given in a basis labelled $(\mathbb{I},\tau)$.

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate connected components 
of average size $O(\log(n))$ and variance $O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected components correspond to sets of anyons that could not have interacted at any 
point in their history. 
We are free to neglect the degrees of freedom associated with braiding between components because each component has trivial total charge.
This allows us to simulate the braiding processes within each component separately. 
In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over components. 
Since each 
component has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are rare 
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

In \Fref{f:syndrome} we show two adjacent sites
involved in a syndrome measurement.
%This will be the total charge on each site, ie. either $\tau$ or $\mathbb{I}.$
The error process has resulted in two pair
creation events along the edge joining these sites.
\dude{i'd like to write out the equations for each of
these 4 steps, but we'd need to keep it tight somehow.}
First, (a) we show the state as a tensor product of two pairs $\tau\tau\otimes\tau\tau$.
Then, (b) these two pairs are embedded into a single larger fusion
tree $\tau\tau\tau\tau$.
This is continuously deformable into the picture in (c).
It is now clear the two braid operations required so
that the observables are represented on the fusion tree.
The general algorithm and data structures used
for this procedure will be discussed elsewhere.

@@

\stf{Cite~\cite{Pfeifer2012} for braiding tricks}

@@

There are several interesting avenues for further research. 
Although it is not completely obvious how to do so, applying these methods to more realistic models such as concrete microscopic spin models or models with non-topological features would give direct insight into practical error rates needed for topological quantum memories in non-Abelian systems. 
It would be extremely interesting to find \emph{classical} spin models whose phase diagram encodes the threshold for error correction in these systems~\cite{Dennis2002}.
%Another potential direction is to incorporate tensor network methods for simulating anyons~\cite{Pfeifer2012, Singh2014} to improve the speed and scale of the simulations. 
Finally, perhaps the most important open question is whether 
\emph{fault-tolerant} decoders can be designed for nonabelian anyon codes.
